---
title: "Earthquakes around the globe"
author: "National Earthquake Information Center (NEIC)"
date: "2023-05-13"
output: html_document
---

ASK: The goal of the project is to determine the Seismic zones that reported earthquake
higher than the magnitude of 7.0

## Installing and required packages

```{r Install Packages, message=FALSE, warning=FALSE}
install.packages("tidyverse",repos = "http://cran.us.r-project.org")
install.packages("lubridate",repos = "http://cran.us.r-project.org")
install.packages("ggthemes",repos= "http://cran.us.r-project.org")
install.packages("ggrepel",repos="http://cran.us.r-project.org")
```

## Loading the packages
```{r Load Packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggrepel)
```
PREPARE: We are given the dataset from the NEIC

## Importin the dataset

```{r Import Dataset, message=FALSE, warning=FALSE}
earthquake <- read.csv("database.csv")
```

PROCESS: The dataset is more or less cleaned, apart from that we may exclude some columns

## Cleaning Dataset

```{r Clean the datasets, message=FALSE, warning=FALSE}
earthquake_1 <- select(earthquake,Date,Time,Latitude,Longitude,Type,Depth,Magnitude)

earthquake_1$Date <- as.Date(earthquake_1$Date)

earthquake_1$Type <- as.factor(earthquake_1$Type)
earthquake_1$Latitude <- round(earthquake_1$Latitude,4)
earthquake_1$Longitude <- round(earthquake_1$Longitude,4)
earthquake_1$Magnitude <- round(earthquake_1$Magnitude,1)
```

ANALYZE: We will analyze the dataset by filtering data with magnitude more than 7.0 and 
tremor caused by only **Earthquakes** and not by **Nuclear Explosions**.

## Analyzing data

```{r Analyzing Data, message=FALSE, warning=FALSE}
earthquake_1 <- earthquake_1 %>% filter(!is.na(earthquake_1$Latitude)) %>% 
  filter(!is.na(earthquake_1$Longitude)) %>% 
  filter(!is.na(earthquake_1$Magnitude))
earthquake_1 <- subset(earthquake_1,Type == "Earthquake" & Magnitude > 7.0)
```

SHARE: We will visualize this data on World map.

## Visualizing Data

```{r Visualize data, message=FALSE, warning=FALSE}
world_coordinates <- map_data("world")

library(ggthemes)
pl <- ggplot() + geom_map(data=world_coordinates,map = world_coordinates,aes(long,lat,map_id=region),color="black",fill="lightblue",size=0.2) +
  geom_point(data=earthquake_1,aes(Longitude,Latitude,size=Magnitude),alpha=0.3,color="red") +
  ggtitle("Significant earthquakes across Globe (1965-2016)",subtitle = "Earthqauke higher than magnitude of 7.0") +
  theme_economist()
Points <- data.frame(lon=c(95.982,142.373),lat=c(3.295,38.297),name=c("Higher than 8.9 Mag","Higher than 8.9 Mag"))
library(ggrepel)
pl2 <- pl + 
  geom_point(data = Points, 
             aes(x = lon, y = lat),
             size = 3) +
  geom_label_repel(data = Points, 
                   aes(x = lon, y = lat, label = name),
                   size = 4,
                   vjust = 0,
                   hjust = 2)
print(pl2)
```

ACT: We can see that most of the earthquake of dangerous magnitude that is higher
than 7.0 ocuured on the **fault line** of the tectonic plates. Hence these areas 
are of great concern, since in the future they can be jolted again by disasterous
earthquakes.
